Imprecise Datapath Design: An Overclocking Approach

### 2013-06-25 TRETS2013/TRETS_overclocking.tex
TRETS modified CSA map
TRETS delete some useless graphs from git repo

We first notice that for all frequency values, the overclocked RCA achieves no larger error expectation than the RCA with truncated operand word-lengths, as predicted by the models for the RCA in Section~\ref{section_RCA_Comparison}. It can also be observed that the CSA cannot be implemented with the original word-length due to the limited area budget and this leads to large truncation errors. Although the CSA with 3 stages is best for some frequencies, the overclocked RCA is still the optimum design when high operating frequencies are applied.

We then perform similar experiments with a variety of area constraints. The optimum design methods with respect to different operating frequencies and area consumptions are demonstrated in Fig.~\ref{Fig_CSA_Tradeoff}. From this figure, several observations can be made. Firstly, if the frequency requirement is not large whilst the available area is large enough to implement a CSA in full precision, it will be the optimum design. This is expected from our earlier analysis in Fig.~\ref{Fig_CSA2stage_Timing}. Secondly, the 2-stage CSA is better than the 3-stage CSA when frequency is low, as it consumes less area although both achieve the same error expectation. Thirdly, only part of the CSA can be implemented under a tighter area budget, whereas the RCA still keeps full precision. In this case, area becomes the dominant factor and precision is lost for the CSA, meaning the RCA with overclocking is the optimum design method across almost the whole frequency domain in this situation. Finally, for very stringent area constraints, the word-length of RCA is also limited. This results in truncation errors initially for all design scenarios. However, the RCA with overclocking can still be employed as the optimum design, because it loses less precision than the CSA under the same area constraint.
%
\begin{figure}[t]
  \centering
  \vspace{-2ex}
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \vspace{-2ex}
  \caption{A demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Fig_CSA_Tradeoff}
\end{figure}

@@

In our experiments, MRE is set ranging from $0.001\%$ to $50\%$. For a certain value of MRE, the design with the maximum operating frequency is selected as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area requirement. Fig.~\ref{Fig_CSA_Tradeoff_Error} can thus be obtained based on these criteria.
%
\begin{figure}[t]
    \centering
    \vspace{-2ex}
    \includegraphics[width=3.7in]{./Figures/Tradeoff_Error.eps}
    \vspace{-2ex}
    \caption{A demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Fig_CSA_Tradeoff_Error}
\end{figure}

@@

In this CCM structure, let the two inputs of the RCA be denoted by $A_S$ and $A_O$ respectively, which are both two's complement numbers. $A_S$ denotes the ``shifted signal", with zeros padded after LSB, while $A_O$ denotes the ``original signal" with MSB sign extension. For an $n$-bit input signal, it should be noted that an $n$-bit RCA is sufficient for this operation, because no carry will be generated or propagated when adding with zeros, as shown in Fig.~\ref{CCM_fig}.
%
\begin{figure}[htbp]
  \vspace{-2ex}
  \centering
  \includegraphics[width=\textwidth]{./Figures/CCM_CarryChain.pdf}
  \vspace{-6ex}
  \caption{Four possible carry chain types in a constant coefficient multiplier with $n$-bit inputs. The notion $s$ denotes the shifted bits and $BP$ denotes the binary point.}
  \label{CCM_fig}
\end{figure}

>>>>

We first notice that for all frequency values, the overclocked RCA achieves no larger error expectation than the RCA with truncated operand word-lengths, as predicted by the models for the RCA in Section~\ref{section_RCA_Comparison}. It can also be observed that the CSA cannot be implemented with the original word-length due to the limited area budget and this leads to large truncation errors. Although the CSA with 4 stages is best for some frequencies, the overclocked RCA is still the optimum design when high operating frequencies are applied.

We then perform similar experiments with a variety of area constraints. The optimum design methods with respect to different operating frequencies and area consumptions are demonstrated in Fig.~\ref{Fig_CSA_Tradeoff}. From this figure, several observations can be made. Firstly, if the frequency requirement is not large whilst the available area is large enough to implement a CSA in full precision, it will be the optimum design. This is expected from our earlier analysis in Fig.~\ref{Fig_CSA2stage_Timing}. Secondly, the 2-stage CSA is better than the 3-stage CSA and the 4-stage CSA when frequency is low, as it consumes less area although all achieve the same error expectation. Thirdly, only part of the CSA can be implemented under a tighter area budget, whereas the RCA still keeps full precision. In this case, area becomes the dominant factor and precision is lost for the CSA, meaning the RCA with overclocking is the optimum design method across almost the whole frequency domain in this situation. Finally, for very stringent area constraints, the word-length of RCA is also limited. This results in truncation errors initially for all design scenarios. However, the RCA with overclocking can still be employed as the optimum design, because it loses less precision than the CSA under the same area constraint.
%
\begin{figure}[t]
  \centering
  \vspace{-2ex}
  \includegraphics[width=4.5in]{./Figures/Tradeoff_Freq.eps}
  \vspace{-2ex}
  \caption{A demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Fig_CSA_Tradeoff}
\end{figure}

@@

In our experiments, MRE is set ranging from $0.001\%$ to $50\%$. For a certain value of MRE, the design with the maximum operating frequency is selected as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area requirement. Fig.~\ref{Fig_CSA_Tradeoff_Error} can thus be obtained based on these criteria.
%
\begin{figure}[t]
    \centering
    \vspace{-2ex}
    \includegraphics[width=4.5in]{./Figures/Tradeoff_Error2.eps}
    \vspace{-2ex}
    \caption{A demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Fig_CSA_Tradeoff_Error}
\end{figure}

@@

In this CCM structure, let the two inputs of the RCA be denoted by $A_S$ and $A_O$ respectively, which are both two's complement numbers. $A_S$ denotes the ``shifted signal", with zeros padded after LSB, while $A_O$ denotes the ``original signal" with MSB sign extension. For an $n$-bit input signal, it should be noted that an $n$-bit RCA is sufficient for this operation, because no carry will be generated or propagated when adding with zeros, as shown in Fig.~\ref{CCM_fig}.
%
\begin{figure}[htbp]
  \vspace{-4ex}
  \centering
  \includegraphics[width=\textwidth]{./Figures/CCM_CarryChain.pdf}
  \vspace{-6ex}
  \caption{Four possible carry chain types in a constant coefficient multiplier with $n$-bit inputs. The notion $s$ denotes the shifted bits and $BP$ denotes the binary point.}
  \label{CCM_fig}
\end{figure}
